R Notebook

### 2020-12-13 AdvendOfCode2020/day_08.Rmd
Commented code from day 08

decodeProgram <- function(.input) {
  tibble(input=.input) %>% 
    separate(input, into = c("cmd","arg"), sep=" ", convert = T) %>% 
    mutate( exec = 0 ) %>% 
    return()
}

@@

  # run the program until before a command be executed a second time
  while( ptr > 0 &&
         ptr <= nrow(.program) &&
         .program[ptr, ]$exec!=1 ) {

@@

input <- readLines("./day08_input.txt")
program <- decodeProgram(input)
result <- executeProgram(program)
changeIndex <- which(program$cmd %in% c("jmp","nop"))
i <- 1

while( result$exit!="end" && i <= length(changeIndex) ){

  step <- program[changeIndex[i],]
  chg.program <- program

  chg.program[changeIndex[i],1] <-  case_when(
    step$cmd=="jmp" ~ "nop",
    step$cmd=="nop" && step$arg!=0 ~ "jmp",
    T ~ chg.program[changeIndex[i],]$cmd
  )

  result <- executeProgram(chg.program)
  i <- i +1
}

result

>>>>

# Parses the array of string into a dataframe with columns "cmd" and "arg"
decodeProgram <- function(.input) {
  tibble(input=.input) %>% 
    separate(input, into = c("cmd","arg"), sep=" ", convert = T) %>% 
    mutate( exec = 0 ) %>% 
    return()
}

@@

  # run the program until..
  while( ptr > 0 &&                # ptr jumps outside the bordering -> its a error
         ptr <= nrow(.program) &&  # ptr jumps outside the bordering -> its a error or finishes
         .program[ptr, ]$exec!=1 ) { # ptr points to a command already executed -> its a loop

@@

# read the input as an array of string
input <- readLines("./day08_input.txt")

# parses the input data
program <- decodeProgram(input)

# execut3 the program as is
result <- executeProgram(program)

# locate the possible changes in the program
changeIndex <- which(program$cmd %in% c("jmp","nop"))
i <- 1

# for each one, change the command and test
while( result$exit!="end" &&        # until we found a change that works
       i <= length(changeIndex) ){  # until the end of possibilities

  # get an step to be changed
  step <- program[changeIndex[i],]

  # modify the program
  chg.program <- program 
  chg.program[changeIndex[i],1] <-  case_when(
    step$cmd=="jmp" ~ "nop",                # change 'jmp' to 'nop'
    step$cmd=="nop" && step$arg!=0 ~ "jmp", # change 'nop' to 'jmp' (when args is not zero: loop)
    T ~ chg.program[changeIndex[i],]$cmd    # not change anything
  )

  # execute the changed program
  result <- executeProgram(chg.program)

  # next change...
  i <- i +1
}

# check the result
result
